---
title: "MA-Analisis"
author: "Carlos López Elorduy"
date: "6/9/2019"
output: html_document
---
```{r}
workingDir <-getwd()
dataDir <- file.path(workingDir, "dades")
resultsDir <- file.path(workingDir, "results")
setwd(resultsDir)
```
Install packages:
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)



if (!require(BiocManager)) install.packages("BiocManager")

installifnot <- function (pkg){
  if (!require(pkg, character.only=T)){
    BiocManager::install(pkg)
}else{
  require(pkg, character.only=T)
  }
}

installifnot("pd.mogene.1.0.st.v1")
installifnot("mogene10sttranscriptcluster.db")
installifnot("oligo")
installifnot("limma")
installifnot("Biobase")
installifnot("arrayQualityMetrics")
installifnot("genefilter")
installifnot("multtest")
installifnot("annotate")
installifnot("xtable")
installifnot("gplots")
installifnot("scatterplot3d")
```



# ---------------------------------------------------------------------------------------------

## LOAD DATA: TARGETS AND CEL FILES. 

# ---------------------------------------------------------------------------------------------

### TARGETS

```{r}
targets <-read.csv(file=file.path(dataDir,"targets.txt"), header = TRUE, sep="\t") 
targets
```

### CELFILES

```{r}
require(GEOquery)
gse<-getGEO("GSE4852")

set<-gse[[1]]
rawData<-set
```


### DEFINE SOME VARIABLES FOR PLOTS

```{r}
sampleNames <- as.character(targets$ShortName)
sampleColor <- as.character(targets$Color)
```


# ---------------------------------------------------------------------------------------------

## QUALITY CONTROL OF ARRAYS: RAW DATA

# ---------------------------------------------------------------------------------------------


### BOXPLOT

```{r}
boxplot(rawData, which="all",las=2, main="Intensity distribution of RAW data", 
        cex.axis=0.6, col=sampleColor, names=sampleNames)
```

### HIERARQUICAL CLUSTERING
```{r}
clust.euclid.average <- hclust(dist(t(exprs(rawData))),method="average")
plot(clust.euclid.average, labels=sampleNames, main="Hierarchical clustering of RawData", 
     cex=0.7,  hang=-1)
```



### PRINCIPAL COMPONENT ANALYSIS
```{r}
plotPCA <- function ( X, labels=NULL, colors=NULL, dataDesc="", scale=FALSE, formapunts=NULL, myCex=0.8,...)
{
  pcX<-prcomp(t(X), scale=scale) # o prcomp(t(X))
  loads<- round(pcX$sdev^2/sum(pcX$sdev^2)*100,1)
  xlab<-c(paste("PC1",loads[1],"%"))
  ylab<-c(paste("PC2",loads[2],"%"))
  if (is.null(colors)) colors=1
  plot(pcX$x[,1:2],xlab=xlab,ylab=ylab, col=colors, pch=formapunts, 
       xlim=c(min(pcX$x[,1])-100000, max(pcX$x[,1])+100000),ylim=c(min(pcX$x[,2])-100000, max(pcX$x[,2])+100000))
  text(pcX$x[,1],pcX$x[,2], labels, pos=3, cex=myCex)
  title(paste("Plot of first 2 PCs for expressions in", dataDesc, sep=" "), cex=0.8)
}

plotPCA(exprs(rawData), labels=sampleNames, dataDesc="raw data", colors=sampleColor,
        formapunts=c(rep(16,4),rep(17,4)), myCex=0.6)
```



### SAVE TO A FILE
```{r}
pdf(file.path(resultsDir, "QCPlots_Raw.pdf"))
boxplot(rawData, which="all",las=2, main="Intensity distribution of RAW data", 
        cex.axis=0.6, col=sampleColor, names=sampleNames)
plot(clust.euclid.average, labels=sampleNames, main="Hierarchical clustering of samples of RawData", 
     cex=0.7,  hang=-1)
plotPCA(exprs(rawData), labels=sampleNames, dataDesc="raw data", colors=sampleColor,
        formapunts=c(rep(16,4),rep(17,4)), myCex=0.6)
dev.off()
```




# ---------------------------------------------------------------------------------------------

## DATA NORMALIZATION

# ---------------------------------------------------------------------------------------------
```{r}
#rawData already normalized
eset<-rawData

write.exprs(eset, file.path(resultsDir, "NormData.txt"))
```




# ---------------------------------------------------------------------------------------------

## QUALITY CONTROL OF ARRAYS: NORMALIZED DATA

# ---------------------------------------------------------------------------------------------

### BOXPLOT
```{r}
boxplot(eset, las=2, main="Intensity distribution of Normalized data", cex.axis=0.6, 
        col=sampleColor, names=sampleNames)
```

### HIERARQUICAL CLUSTERING
```{r}
clust.euclid.average <- hclust(dist(t(exprs(eset))),method="average")
plot(clust.euclid.average, labels=sampleNames, main="Hierarchical clustering of NormData", 
     cex=0.7,  hang=-1)
```


### PRINCIPAL COMPONENT ANALYSIS
```{r}
plotPCA <- function ( X, labels=NULL, colors=NULL, dataDesc="", scale=FALSE, formapunts=NULL, myCex=0.8,...)
{
  pcX<-prcomp(t(X), scale=scale) # o prcomp(t(X))
  loads<- round(pcX$sdev^2/sum(pcX$sdev^2)*100,1)
  xlab<-c(paste("PC1",loads[1],"%"))
  ylab<-c(paste("PC2",loads[2],"%"))
  if (is.null(colors)) colors=1
  plot(pcX$x[,1:2],xlab=xlab,ylab=ylab, col=colors, pch=formapunts, 
       xlim=c(min(pcX$x[,1])-10, max(pcX$x[,1])+10),ylim=c(min(pcX$x[,2])-10, max(pcX$x[,2])+10))
  text(pcX$x[,1],pcX$x[,2], labels, pos=3, cex=myCex)
  title(paste("Plot of first 2 PCs for expressions in", dataDesc, sep=" "), cex=0.8)
}

plotPCA(exprs(eset), labels=sampleNames, dataDesc="NormData", colors=sampleColor,
        formapunts=c(rep(16,4),rep(17,4)), myCex=0.6)
```



### SAVE TO A FILE
```{r}
pdf(file.path(resultsDir, "QCPlots_Norm.pdf"))
boxplot(eset, las=2, main="Intensity distribution of Normalized data", cex.axis=0.6, 
        col=sampleColor, names=sampleNames)
plot(clust.euclid.average, labels=sampleNames, main="Hierarchical clustering of NormData", 
     cex=0.7,  hang=-1)
plotPCA(exprs(eset), labels=sampleNames, dataDesc="selected samples", colors=sampleColor,
        formapunts=c(rep(16,4),rep(17,4)), myCex=0.6)
dev.off()
```



### ARRAY QUALITY METRICS

```{r}
arrayQualityMetrics(eset,  reporttitle="QualityControl", force=TRUE)
```


# ---------------------------------------------------------------------------------------------

## FILTER OUT THE DATA
# ---------------------------------------------------------------------------------------------
```{r}
annotation(eset) <- "org.Mm.eg.db"
eset_filtered <- nsFilter(eset, var.func=IQR,
         var.cutoff=0.75, var.filter=TRUE,
         filterByQuantile=TRUE)
```


### NUMBER OF GENES OUT
```{r}
print(eset_filtered$filter.log$numLowVar)
```



### NUMBER OF GENES IN
```{r}
print(eset_filtered$eset)
```




# ---------------------------------------------------------------------------------------------

## DIFERENTIAL EXPRESSED GENES SELECTION. LINEAR MODELS. COMPARITIONS

# ---------------------------------------------------------------------------------------------

### CONTRAST MATRIX.lINEAR MODEL
```{r}
treat <- targets$Experiment
lev <- factor(treat, levels = unique(treat))
design <-model.matrix(~0+lev)
colnames(design) <- levels(lev)
rownames(design) <- sampleNames
print(design)
```



### COMPARISON
```{r}
cont.matrix1 <- makeContrasts( 
        Control.vs.Nickel = HUVEC_control_experiment-HUVEC_nickel_experiment,
        levels = design)
comparison1 <- "Effect of Induction"
```


### MODEL FIT
```{r}
fit1 <- lmFit(eset_filtered$eset, design)
fit.main1 <- contrasts.fit(fit1, cont.matrix1)
fit.main1 <- eBayes(fit.main1)
```




# ---------------------------------------------------------------------------------------------

## DIFERENTIAL EXPRESSED GENES LISTS.TOPTABLES

# ---------------------------------------------------------------------------------------------

### FILTER BY FALSE DISCOVERY RATE AND FOLD CHANGE
```{r}
topTab <-  topTable (fit.main1, number=nrow(fit.main1), coef="Control.vs.Nickel", adjust="fdr",lfc=abs(3))
```



### EXPORTED TO CSV AND HTML FILE
```{r}
write.csv2(topTab, file= file.path(resultsDir,paste("Selected.Genes.in.comparison.",
                                                    comparison1, ".csv", sep = "")))
print(xtable(topTab,align="lllllllllllllllllllllll"),type="html",html.table.attributes="",
      file=paste("Selected.Genes.in.comparison.",comparison1,".html", sep=""))

```



# ---------------------------------------------------------------------------------------------

## VOLCANO PLOTS

# ---------------------------------------------------------------------------------------------
```{r}

volcanoplot(fit.main1, highlight=10, names=fit.main1$ID, 
            main = paste("Differentially expressed genes", colnames(cont.matrix1), sep="\n"))
abline(v = c(-3, 3))


pdf(file.path(resultsDir,"Volcanos.pdf"))
volcanoplot(fit.main1, highlight = 10, names = fit.main1$ID, 
            main = paste("Differentially expressed genes", colnames(cont.matrix1), sep = "\n"))
abline(v = c(-3, 3))
dev.off()

```


# ---------------------------------------------------------------------------------------------

## DATA ANNOTATION

# ---------------------------------------------------------------------------------------------
```{r}

all_anota<-data.frame(exprs(eset))
Annot <- data.frame(SYMBOL=sapply(contents(mogene10sttranscriptclusterSYMBOL), paste, collapse=", "),
                    DESC=sapply(contents(mogene10sttranscriptclusterGENENAME), paste, collapse=", "))
Annot<-Annot[!Annot$SYMBOL=="NA",]
Annot<-Annot[!Annot$DESC=="NA",]
head(Annot)

anotaGenes <- merge(Annot,all_anota, by.x=0,by.y=0)
head(anotaGenes)
write.table(anotaGenes, file ="data.ann.txt",sep="\t")

rownames(anotaGenes) <- anotaGenes[,1]
anotaGenes <- anotaGenes[,-1]
anotaGenes.end <- merge(anotaGenes, topTab, by.x=0,by.y=0)

topTab.end <- anotaGenes.end[,c(1:3,12:17,4:11)]

rownames(topTab.end) <- topTab.end[,1]
topTab.end <- topTab.end[, -1]
write.csv(topTab.end, file = file.path(resultsDir,"TopTable.end.csv"))
```

