---
title: "**Exercise 1**"
author: "Carlos López Elorduy"
date: "5/15/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction
**Title:** Gene expression profile of nickel-stimulated HUVEC

**Organism:** Homo sapiens

**Experiment type:** Expression profiling by array.

**Status:** Public on Dec 31, 2006

## Summary
Nickel compounds, important carcinogens and prime inducers of contact allergy in humans, are potent activators of endothelial cells (ECs). To elucidate the signaling pathways mediating the cellular responses to nickel we performed Affymetrix gene profiling of human primary ECs.
Keywords: stress response

## Experimental design
HUVEC cultured under standard conditions were treated for 5h with 1.5 mM NiCl2 or left untreated before total RNA isolation and processing for microarray hybridization. Four independent experiments were performed, i.e. 8 samples (Control and experiment) were analyzed.

## Citations
    Viemann D, Schmidt M, Tenbrock K, Schmid S et al. The contact allergen nickel triggers a unique inflammatory and proangiogenic gene expression pattern via activation of NF-kappaB and hypoxia-inducible factor-1alpha. J Immunol 2007 Mar 1;178(5):3198-207. PMID: 17312168

## Data description
We have two files:

  - The **targets** file
  - The **expression matrix** file

These files can be either imported manually using the Rstudio tools or using the following commands:
```{r}
targets<-read.csv("targets.txt",header = T, sep = "\t")
myExpress<-as.matrix(read.table("expression.txt", header=T, sep="\t", row.names=1, as.is=T))
```

### Targets
```{r}
show(targets)
```

### Expression matrix
```{r}
dim(myExpress)
summary(myExpress)
```

## ExpressionSet creation

The expression set will contain all the relevant information about our study, starting by the expression matrix:
```{r}
library("Biobase")
myEset<-ExpressionSet(myExpress)
```

The next information we have to add is the information stored in the targets.
```{r}
columnDesc<-data.frame(labelDescription=c("SampleName", "Experiment", "Replicate", "Color"), row.names = c("SampleName", "Experiment", "Replicate", "Color"))
myAnnotDF <- new("AnnotatedDataFrame", data=targets, varMetadata= columnDesc)
targets
columnDesc

```
We create an array of the length of the total number of genes (2283 in my case)
```{r}
myGenes <-  paste0("gene",1:2283)
```


```{r}
rownames(pData(myAnnotDF))<-pData(myAnnotDF)$SampleName
myEset <- ExpressionSet(assayData=myExpress, phenoData=myAnnotDF)
```

We finally add the array of genes to the expression set:
```{r}
myEset <- ExpressionSet(assayData=myExpress,featureNames =myGenes)
show(myEset)
```

I can add also my contact info to the expression set:
```{r}
myInfo=list(myName="Carlos Lopez Elorduy", myLab="Bioinformatics Lab",myContact="carlos.lopez@alum.esci.upf.edu", myTitle="Practical Exercise on ExpressionSets")
```
